---
html_document:
  df_print: paged
author: "Corinne Medeiros"
date: "September 20th 2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
title: "Improving a Dodgers Marketing Promotion - Modeling in R"
word_document: default
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```
### **Objective:** Determine what night would be the best to run a marketing promotion to increase attendance.

**Data source:**  
Dodgers Major League Baseball data from 2012  
[Data Source: dodgers.csv](https://corinnemedeiros.github.io/Projects/Dodgers-Marketing/dodgers.csv)

After using Python to explore the Dodgers data and create some visualizations, I am now using R to create Regression models. Looking at the model summary results will help to determine which factors might be the best predictors of attendance.

```{r echo = FALSE, message = FALSE}
library(boot)
library(ggm)
library(ggplot2)
library(Hmisc)
library(polycor)
library(readxl)
library(car)
library(QuantPsyc)
library(broom)

# Loading data
dodgers_data <- readr::read_csv("dodgers.csv")

```

```{r}
# All variables in dataset
colnames(dodgers_data)

```

### **Splitting data into train and test sets**  

```{r}
library(caTools)

# Ensuring random numbers are generated
set.seed(123) 

# Splitting data into 80/20 ratio
sample = sample.split(dodgers_data, SplitRatio = 0.20)

# Creating a training dataset and testing dataset
dodgers_train = subset(dodgers_data, sample==FALSE)
dodgers_test = subset(dodgers_data, sample==TRUE)

```
### **Regression models using training data**  

```{r}
# Regression model using 'date'
dodgers_data_date_mod <- lm(attend ~ day, data = dodgers_train)
summary(dodgers_data_date_mod)
```

```{r}
# Regression model using 'day of week'
dodgers_data_dow_model <- lm(attend ~ day_of_week, data = dodgers_train)
summary(dodgers_data_dow_model)

```

```{r}
# Regression model using 'day of week' and 'month'
dodgers_data_dow_month_mod <- lm(attend ~ day_of_week + month, data = dodgers_train)
summary(dodgers_data_dow_month_mod)

```

### **Comparing performance of models**  

```{r}
# Metrics for model 1 - 'date'
glance(dodgers_data_date_mod) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)

```

```{r}
# Metrics for model 2 - 'day of week'
glance(dodgers_data_dow_model) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)

```

```{r}
# Metrics for model 3 - 'day of week' and 'month'
glance(dodgers_data_dow_month_mod) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)

```

### **Regression model using testing data**  

```{r}
# Regression model using 'day of week' and 'month'
dodgers_data_dow_month_mod <- lm(attend ~ day_of_week + month, data = dodgers_test)
summary(dodgers_data_dow_month_mod)

```

```{r}
# Metrics for model 3 - 'day of week' and 'month' (using testing data set)
glance(dodgers_data_dow_month_mod) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)

```


### **Conclusion:**

Based on the summaries of these regression models, month and day of the week are the best predictors of attendance at a Dodgers game in 2012. The model using day of the week and month (model 3) had the lowest AIC and BIC scores, the highest R2, and the lowest F-statistic p.value, making it more statistically significant. Running the same model using the testing data also proved successful, with a high R2 and low p.value. All in all, in order to improve attendance, it would be most effective to run a marketing promotion on **Tuesdays in June.** This way, you're most likely to reach the largest audience for the promotion. 
